---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE}
knitr::opts_chunk$set(error = TRUE)
```

```{r}
#load libraries
library(ggplot2)
library(plyr)
library(dplyr)

#load data from CSV file
activity_data <- read.csv('C:/Users/smatismi/OneDrive - Capgemini/Documents/R/win-library/3.5/rmarkdown/activity.csv')

str(activity_data)
```

```{r} 

steps_day <- tapply(activity_data$steps, activity_data$date, sum, na.rm=TRUE)
steps_day <- steps_day[!is.na(steps_day)]

str(steps_day)

```

```{r}
hist(x=steps_day, col="green", breaks = 20, main= "Steps Per Day", xlab = "Total steps", ylab ="Frequency")
#cleansteps_day <-steps_day[!is.na(steps_day$steps,)]
```

## What is mean total number of steps taken per day?

```{r}
mean_st <- mean(steps_day)
mean_st
```

```{r}
med_st <-median(steps_day)
med_st
```

## What is the average daily activity pattern?
```{r}
#cleansteps_day <-steps_day[!is.na(steps_day$steps),]
#intervalTb <- steps_day[ ,c (lapply(.SD, mean, na.rm=TRUE)), .SDcols =c("steps"), by =.(interval)]
#intervalTb <- ddply(steps_day, .(interval), summarize, Ave = mean(steps))
interval <- activity_data %>%
  filter(!is.na(steps)) %>%
  group_by(interval) %>%
  summarize(Avg_steps = mean(steps))
  
ggplot(interval, aes(x=interval , y=Avg_steps)) + geom_line(color='green', size=1) + labs(title ="Average Steps per Day", x ="interval", y = "average steps per day" )
```

#Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
#interval[steps == max(steps), .(max_interval <- interval)]
max_steps <- max(interval$Avg_steps)
interval[interval$Avg_steps == max_steps, ]
```


## Imputing missing values

##imputing missing values, calculate the total number of missing values
```{r}
nrow(activity_data[is.na(activity_data$steps),])
```

#Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.

#substute missing values with the mean
```{r}
activity_data_fill <- activity_data
nmv <- is.na(activity_data_fill$steps)
step_ave <- tapply(activity_data_fill$steps, activity_data_fill$interval, mean, na.rm=TRUE)
activity_data_fill$steps[nmv] <- step_ave[as.character(activity_data_fill$interval[nmv])]

new_activity <- tapply(activity_data_fill$steps, activity_data_fill$date, sum, na.rm = TRUE)
hist(x=new_activity, col="green", breaks=20, xlab="Daily Steps", ylab="Frequency", main="Steps per Day missing values filled with mean")
```

```{r}
mean_fill <- mean(new_activity)
mean_fill
```

```{r}
median_fill <- median(new_activity)
median_fill
```

## Are there differences in activity patterns between weekdays and weekends?
##Are there differences in activity patterns between weekdays and weekends?
```{r}
#is_wk_day <- function(d) {
 # wd <-weekdays(d)
#  ifelse (wd == "Saturday" | wd == "Sunday", #"weekend" , "weekday")
#}

#we <- sapply(activity_data_fill$date, is_wk_day)
#activity_data$wk <-  as.factor(we)
#head(activity_data_fill)
activity_data_fill$realdate <- as.Date(activity_data_fill$date, format = "%Y-%m-%d")
activity_data_fill$weekday <- weekdays(activity_data_fill$realdate)
activity_data_fill$weekorend <- ifelse(activity_data_fill$weekday =="Saturday" | activity_data_fill$weekday =="Sunday", 'weekend', 'weekday' )
head(activity_data_fill)
```

```{r}
steptable <- aggregate(steps~interval+weekorend, data=activity_data_fill, FUN = mean)
head(steptable)

```

```{r}
steptable$time <- steptable$interval/100
head(steptable)
J<-ggplot(steptable, aes(time, steps))
J+geom_line(col="green") + ggtitle("Average Steps per 5 Minute Interval, Weekday verses Weekend") +xlab("time")+ylab("steps")+facet_wrap(~weekorend, ncol= 1, nrow =2)
```
